---
title: "Using Icons in Figures with {ggsvg}"
author:
  - name: "Haley Epperly Fox"
    orcid: "0000-0001-6102-4784"

abstract: "Learn how to insert icons into your plots using the `{ggsvg}` package."

date: Sep 5, 2022
date-format: MMMM D, YYYY

image: ../images/ggsvg-penguin.svg

categories: graph aesthetics
bibliography: references.bib
---

------------------------------------------------------------------------

Have you ever seen a figure with cute little icons representing data points and thought how you would love to recreate that? I'm here to tell you that you can and the `ggsvg` package makes it super simple.

All you have to do is:

1.  Load in the `ggsvg` package

2.  Figure out which icon you want to use from the ones available [here](https://www.svgrepo.com/)

3.  Save the url for your chosen icon to read into your gpplot code

4.  Use `geom_point_svg` in your ggplot code referencing your saved url [@ggsvg]

Let's run through an example with the `palmerpenguins` dataset [@palmerpenguins]. Checkout the package [website](https://allisonhorst.github.io/palmerpenguins/) for some more info and fun artwork! We're going to subset these data to only include female Chinstrap penguins so that we have fewer data points to plot.

::: callout-note
#### Example

```{r}
#| label: load-libraries
#| output: false

# load libraries
library(palmerpenguins)
library(tidyverse)
# install.packages('rsvg')
# remotes::install_github('coolbutuseless/ggsvg')
library(rsvg)
library(ggsvg)
```

```{r}
#| label: plot

# find the icon you want and save the url for later reference following this format
svg_url <- 'https://www.svgrepo.com/download/133788/penguin.svg'
svg_txt <- paste(readLines(svg_url), collapse = "\n")

# filter the penguin data to only female Chinstrap penguins
chinstrap <- penguins %>% 
  filter(species == "Chinstrap", sex == "female")

# plot the data with the selected icon used instead of points
ggplot(chinstrap) + 
  geom_point_svg(aes(body_mass_g, flipper_length_mm), svg = svg_txt) + 
  labs(x = "Body Mass (g)", y = "Flipper Length (mm)", 
       title = "Comparing body mass and flipper length of female Chinstrap penguins") +
  theme_classic()
```
:::
